---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# eesyscreener <a href="https://dfe-analytical-services.github.io/eesyscreener/"><img src="" align="right" height="138" /></a>

<!-- badges: start -->
[![R-CMD-check](https://github.com/dfe-analytical-services/eesyscreener/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/dfe-analytical-services/eesyscreener/actions/workflows/R-CMD-check.yaml)
[![pkgdown](https://github.com/dfe-analytical-services/eesyscreener/actions/workflows/pkgdown.yaml/badge.svg)](https://github.com/dfe-analytical-services/eesyscreener/actions/workflows/pkgdown.yaml)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

This is a package designed to provide documentation and reusable code supprting the data standards for the explore education statistics (EES) platform.

The package contains:

- A core `screen_csv()` function, built from `screen_dfs()`, and `screen_filenames()` and the constituent individual checks, as well as a related wrapper function `screen_zip()`
- Functions to generate test data
- Example datasets to aid testing and demonstration
- Additional functions to aid in screening / preparing data for EES

There are three main use cases:

- Providing the logic behind the [data screener R Shiny app](https://github.com/dfe-analytical-services/dfe-published-data-qa), used to check files before upload
- Using that logic within EES itself through a [plumber API implementation](https://github.com/dfe-analytical-services/ees-screener-api), screening uploads into the service directly
- Analysts using the R functions within their own pipelines as pre-emptive quality assurance

## Installation

eesyscreener is not currently available on CRAN. For the time being you can install the development version from GitHub.

``` r
# install.packages("pak")
pak::pak("dfe-analytical-services/eesyscreener")
```

## Minimal example

This shows a quick reproducible example you can run in the console to test with. It also shows an example of the output structure from the core `screen_csv()` function.
```{r}
library(eesyscreener)

# Create example temporary CSV files
data_file <- tempfile(fileext = ".csv")
meta_file <- tempfile(fileext = ".meta.csv")

data.table::fwrite(example_data, data_file)
data.table::fwrite(example_meta, meta_file)

result <- screen_csv(
  data_file,
  meta_file,
  "data.csv",
  "data.meta.csv"
)

result$results_table |>
  head()

result$overall_stage

result$overall_message

# Clean up temporary CSV files
file.remove(data_file)
file.remove(meta_file)
```

## Example CSVs

Quick examples of how to make use of the data within the package to generate CSVs for testing:
``` r
library(eesyscreener)

write.csv(example_data, "example_data.csv", row.names = FALSE)
write.csv(example_meta, "example_data.meta.csv", row.names = FALSE)

# Generate a file pairing that will fail the tests (spaces in filename)
write.csv(example_data, "example data.csv", row.names = FALSE)
write.csv(example_meta, "example data.meta.csv", row.names = FALSE)
```

Other available example files can be found on the documentation site under examples. Use `write.csv()` as in the examples above to generate CSVs from them.

## Generate big test files

If you want to generate larger files for testing with, you can use the `generate_test_dfs()` function to create files with any number of time periods, locations, filters and indicators.

```r
files <- eesyscreener::generate_test_dfs(
  years = 2013:2015, 
  pcon_names = "Sheffield Central", 
  pcon_codes = "E14000919", 
  num_filters = 2, 
  num_indicators = 3
)

# Data and metadata are returned in a list, to extract:
df <- files$data
df_meta <- files$meta
```

If you want to go really big, combine with the [dfeR package](https://dfe-analytical-services.github.io/dfeR/), to pass in vectors of Parliamentary Constituencies, and then [data.table](https://rdatatable.gitlab.io/data.table/) for much faster CSV creation.

The following example creates an example data and metadata pair with a data set of just over 6 million rows. Formula to calculate rows is:

- length(years) * length(pcon_codes) * (5 ^ num_filters)

```r
# Load this eesyscreener package
devtools::load_all()

# Additional dependencies
library("dfeR")
library("data.table")

# Get a data frame of all Parliamentary Constituencies in England
pcons <- dfeR::fetch_pcons(countries = "England")

beefy <- eesyscreener::generate_test_dfs(
  years = c(1980:2025),
  pcon_codes = pcons$pcon_code,
  pcon_names = pcons$pcon_name,
  num_filters = 3,
  num_indicators = 45,
  verbose = TRUE
)

# Then to create CSVs, use data.table as it's much faster
data.table::fwrite(beefy$data, "beefy_data.csv")
data.table::fwrite(beefy$meta, "beefy_data.meta.csv")
```

------------------------------------------------------------------------

## Contributing

Ideas for eesyscreener should first be raised as a [GitHub
issue](https://github.com/dfe-analytical-services/eesyscreener/issues/new/choose) after which
anyone is free to write the code and create a pull request for review.

For more details on contributing to eesyscreener, see our [contributing
guidelines](https://dfe-analytical-services.github.io/eesyscreener/CONTRIBUTING.html).
